---
title: "3CT"
output:
  pdf_document: default
  html_document: default
date: '2022-12-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}
setwd("~/Documents/TDP control files/Control study/")
# Loading
library(pacman)
p_load(readxl, tidyverse)
p_load(broom, ggsignif, multcomp, emmeans, DescTools)
# xlsx files
df <- read_excel("3CT_controls.xlsx")


```


Assign genotypes
```{r}
df$Genotype <- 0


df$Genotype <- ifelse(df$ID==82, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==84, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==107, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==109, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==80, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==86, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==115, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==117, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==74, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==76, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==119, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==121, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==129, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==137, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==141, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==98, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==102, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==112, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==120, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==124, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==128, 'NEFH-tTAxTDP', df$Genotype)

df$Genotype <- factor(df$Genotype, levels = c("WT", "NEFH-tTA", "TDP", "NEFH-tTAxTDP"))

```

We will primarily analyse time spent in the different zones. We expect:

Habituation phase: time spent in neutral, social, and novel zone will be equal.
For social phase (1 mouse introduced): time spent in social will be bigger than neutral and novel zone combined.
for novelty phase (1 novel mouse introduced into novelty zone): time spent in novelty zone will be bigger then in neutral and social zone combined.


Divide dataset up into Habituation, social, and novelty phases

```{r}
habituation_cond <-  df %>% filter(Condition == "Hab")
social_cond<-  df %>% filter(Condition == "Social")
novelty_cond<-  df %>% filter(Condition == "Novel")
```


```{r}

habituation_cond <- habituation_cond %>% mutate(social_preference_relative=((`In zone SocialZone cumulative s`-`In zone NoveltyZone cumulative s`)/(`In zone NoveltyZone cumulative s`+`In zone SocialZone cumulative s`)))


#As described in Baronio 2015 
#https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0116363&type=printable 

social_cond <- social_cond %>% mutate(social_preference_relative=((`In zone SocialZone cumulative s`-`In zone NoveltyZone cumulative s`)/(`In zone NoveltyZone cumulative s`+`In zone SocialZone cumulative s`)))

novelty_cond <- novelty_cond %>% mutate(novelty_preference_relative=((`In zone NoveltyZone cumulative s`-`In zone SocialZone cumulative s`)/(`In zone SocialZone cumulative s`+`In zone NoveltyZone cumulative s`)))

```



testing by genotype:

Dunnet tests

```{r}
###Relative preference
anova_social_preference_relative <- aov(social_preference_relative ~ Genotype, data = social_cond)
summary(anova_social_preference_relative)

#Dunnet's post hoc
emmeans(anova_social_preference_relative ,specs = trt.vs.ctrl~ Genotype)
```


```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```

Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```

```{r}

Dunnet<- glht(anova_social_preference_relative, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))


ggplot(data= social_cond, aes(x=Genotype, y=social_preference_relative)) +
  ylab("Sociability Index")+
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Times New Roman"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(1.2, 1.1,1)
  )
```
Social novelty

```{r}

###Relative preference
anova_novel_preference_relative <- aov(novelty_preference_relative ~ Genotype, data = novelty_cond)
summary(anova_novel_preference_relative)

#Dunnet's post hoc
DunnettTest(x=novelty_cond$novelty_preference_relative, g=novelty_cond$Genotype)

```


```{r}

Dunnet<- glht(anova_novel_preference_relative, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))


ggplot(data= novelty_cond, aes(x=Genotype, y=novelty_preference_relative)) +
  ylab("Social Novelty Preference Index")+
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Times New Roman"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(0.65,0.55,0.45)
  )
```

Tukey tests

Test genotype differences - social preference


#assumption tests

```{r}
p_load(car)

#Independence
#Should be okay

#outliers - 3, one is extreme
p_load(rstatix)
outliers <- social_cond %>%
  group_by(Genotype) %>%
  identify_outliers(social_preference_relative)

#normality - TDP distribution significant
by(social_cond$social_preference_relative, social_cond$Genotype, shapiro.test)

#Homogeneity of variance - holds
leveneTest(social_cond$social_preference_relative, social_cond$Genotype, center = mean)
bartlett.test(social_preference_relative ~Genotype, data=social_cond)

kruskal.test(social_preference_relative ~ Genotype, 
             data = social_cond)

dunnTest(social_preference_relative ~ Genotype, 
             data = social_cond,
         method="bonferroni")
```

```{r}

###Relative preference
anova_social_preference_relative <- aov(social_preference_relative ~ Genotype, data = social_cond)
summary(anova_social_preference_relative)

#Dunnet's post hoc
emmeans(anova_social_preference_relative ,specs = trt.vs.ctrl~ Genotype)

#Tukey
TukeyHSD(anova_social_preference_relative, conf.level = 0.95)

emm<-emmeans(anova_social_preference_relative, pairwise ~ Genotype, conf.level = 0.95)
emm
s_emm<- as.data.frame(test(emm)[2])

Tukey_social <- TukeyHSD(anova_social_preference_relative, conf.level = 0.95)
s_df<- as.data.frame(Tukey_social[1])
colnames(s_df) <- c('Mean difference','CI lower','CI upper', 'Adjusted p value')

s_emm
s_df
s_df$SE <- s_emm$contrasts.SE
s_df <- s_df[, c("Mean difference", "SE",'CI lower','CI upper', 'Adjusted p value')]

```

Make nice table of results
```{r}
#caption = "<center><strong>CENTER</strong></center>"
p_load(knitr, kableExtra)
s_df %>% kable("html", digits=4,caption = "<center>Sociability Index</center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, html_font = "Times New Roman" )


```


```{r}

###Relative preference
anova_social_preference_relative <- aov(social_preference_relative ~ Genotype, data = social_cond)
summary(anova_social_preference_relative)

#Tukey
TukeyHSD(anova_social_preference_relative, conf.level = 0.95)

emm<-emmeans(anova_social_preference_relative, pairwise ~ Genotype, conf.level = 0.95)
emm
s_emm<- as.data.frame(test(emm)[2])

Tukey_social <- TukeyHSD(anova_social_preference_relative, conf.level = 0.95)
s_df<- as.data.frame(Tukey_social[1])
colnames(s_df) <- c('Mean difference','CI lower','CI upper', 'Adjusted p value')

s_emm
s_df
s_df$SE <- s_emm$contrasts.SE
s_df <- s_df[, c("Mean difference", "SE",'CI lower','CI upper', 'Adjusted p value')]

```

Make nice table of results
```{r}
#caption = "<center><strong>CENTER</strong></center>"
p_load(knitr, kableExtra)
s_df %>% kable("html", digits=4,caption = "<center>Sociability Index</center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, html_font = "Times New Roman" )


```


##Plots


Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice



```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```

Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```

```{r}


Dunnet<- glht(anova_social_preference_relative, mcp(Genotype = "Tukey")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= social_cond, aes(x=Genotype, y=social_preference_relative)) +
  ylab("Sociability Index")+
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Times New Roman"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(1.5,1.4,1.3,1.2, 1.1,1)
  )
```



Test genotype differences - novelty preference


#assumption tests

```{r}
p_load(car)

#Independence
#Should be okay

#outliers - 0
p_load(rstatix)
outliers <- novelty_cond %>%
  group_by(Genotype) %>%
  identify_outliers(novelty_preference_relative)

#normality - holds
by(novelty_cond$novelty_preference_relative, social_cond$Genotype, shapiro.test)

#Homogeneity of variance - holds
leveneTest(novelty_cond$novelty_preference_relative, novelty_cond$Genotype, center = mean)
bartlett.test(novelty_preference_relative ~Genotype, data=novelty_cond)

#kruskal.test(novelty_preference_relative ~ Genotype, 
         #    data = novelty_cond)
```


```{r}

###Relative preference
anova_novel_preference_relative <- aov(novelty_preference_relative ~ Genotype, data = novelty_cond)
summary(anova_novel_preference_relative)

#Dunnet's post hoc
emmeans(anova_novel_preference_relative ,specs = trt.vs.ctrl~ Genotype)


#Tukey
TukeyHSD(anova_novel_preference_relative, conf.level = 0.95)

emm<-emmeans(anova_novel_preference_relative, pairwise ~ Genotype, conf.level = 0.95)
emm
sn_emm<- as.data.frame(test(emm)[2])

Tukey_social_novelty <- TukeyHSD(anova_novel_preference_relative, conf.level = 0.95)
sn_df<- as.data.frame(Tukey_social_novelty[1])
colnames(sn_df) <- c('Mean difference','CI lower','CI upper', 'Adjusted p value')

sn_emm
sn_df
sn_df$SE <- sn_emm$contrasts.SE
sn_df <- sn_df[, c("Mean difference", "SE",'CI lower','CI upper', 'Adjusted p value')]

```


Make nice table of results
```{r}
p_load(knitr, kableExtra)
sn_df %>% kable("html", digits=4, caption = "<center>Social Novelty Preference Index</center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, html_font = "Times New Roman")

```


```{r}

Dunnet<- glht(anova_novel_preference_relative, mcp(Genotype = "Tukey")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= novelty_cond, aes(x=Genotype, y=novelty_preference_relative)) +
  ylab("Social Novelty Preference Index")+
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Times New Roman"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(1.5,1.4,1.3,1.2, 1.1,1)
  )
```
