---
title: "LDCT"
output:
  pdf_document: default
  html_document: default
  word_document: default
date: '2022-12-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load data
```{r}
setwd("~/Documents/TDP control files/Control study/")
# Loading
library(pacman)
p_load(readxl, tidyverse, DescTools)
# xlsx files
df <- read_excel("EPM_controls.xlsx")


```


Assign genotypes
```{r}
df$Genotype <- 0



df$Genotype <- ifelse(df$ID==82, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==84, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==107, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==109, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==80, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==86, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==115, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==117, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==74, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==76, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==119, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==121, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==129, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==137, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==141, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==98, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==102, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==112, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==120, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==124, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==128, 'NEFH-tTAxTDP', df$Genotype)

df$Genotype <- factor(df$Genotype, levels = c("WT", "NEFH-tTA", "TDP", "NEFH-tTAxTDP"))


```

Time spent in dark chamber:
```{r}
#periphery_preference = time spent in periphery / time spent in center
 df<-  df %>% mutate(time_dark=df$`In zone Closed_arm1 cumulative`+df$`In zone Closed_arm2 cumulative`)
 df<-  df %>% mutate(time_light=df$`In zone Open_arm1 cumulative`+df$`In zone Open_arm2 cumaltive`)
# 
# #relative time spent in dark over light
 df<-  df %>% mutate(light_preference=time_light/(time_light+time_dark)*100) 


```

Frequency in arms
```{r}
#frequency in dark and light arms
df<-  df %>% mutate(freq_dark=df$`In zone Closed_arm1  frequency`+df$`In zone Closed_arm2 frequency`)
df<-  df %>% mutate(freq_light=df$`In zone Open_arm1 frequency`+df$`In zone Open_arm2 frequency`)

#relative time spent in dark over light as entries in open arm / total arm entries
df<-  df %>% mutate(light_preference_freq=freq_light/(freq_light+freq_dark)*100) 

```

testing by genotype:

Dunnet

comparing relative darkness preference FREQUENCY
```{r}

anova_light_pref <- aov(light_preference_freq ~ Genotype, data = df) #if using light_preference_freq instead, it is non-significant
summary(anova_light_pref)

p_load(DescTools)
DunnettTest(x=df$light_preference_freq, df$Genotype)
```

comparing relative darkness preference TIME
```{r}

anova_light_pref_t <- aov(light_preference ~ Genotype, data = df) #if using light_preference_freq instead, it is non-significant
summary(anova_light_pref_t)

DunnettTest(x=df$light_preference, df$Genotype)

```

Plots

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice
```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```


Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}



```

lightpreference time
```{r}
p_load(multcomp, broom, ggsignif)

Dunnet<- glht(anova_light_pref_t, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df, aes(x=Genotype, y=light_preference)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  theme(text = element_text(family = "Times New Roman"))+
  ylab("Open Arm Time Percentage")+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(51,49,47,45,43,41)
  )

```
lightpreference frequency
```{r}
p_load(broom, ggsignif, multcomp)

Dunnet<- glht(anova_light_pref, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df, aes(x=Genotype, y=light_preference_freq)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  theme(text = element_text(family = "Times New Roman"))+
  ylab("Open Arm Entry Percentage")+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(71,69,67,65,63,61)
  )

```


Tukey

comparing relative darkness preference TIME

#assumption tests

```{r}
p_load(car)

#Independence
#Should be okay

#outliers- 0
p_load(rstatix)
outliers <- df %>%
  group_by(Genotype) %>%
  identify_outliers(light_preference)

#normality - holds
by(df$light_preference, df$Genotype, shapiro.test)

#Homogeneity of variance - does not hold
leveneTest(df$light_preference, df$Genotype, center = mean)
bartlett.test(df$light_preference ~df$Genotype, data=df)

kruskal.test(light_preference ~ Genotype, 
             data = df)

p_load(pgirmess)

kruskalmc(light_preference ~ Genotype, 
             data = df)
p_load(FSA)

dunnTest(light_preference ~ Genotype, 
             data = df,
         method="bonferroni")
```


```{r}
anova_light_pref_t <- aov(light_preference ~ Genotype, data = df) #if using light_preference_freq instead, it is non-significant
summary(anova_light_pref_t)

#p_load(emmeans)
p_load(emmeans)
emm<-emmeans(anova_light_pref_t, pairwise ~ Genotype, conf.level = 0.95)
emm
t_emm<- as.data.frame(test(emm)[2])

Tukey_light <- TukeyHSD(anova_light_pref_t, conf.level = 0.95)
Tukey_light

t_df<- as.data.frame(Tukey_light[1])
colnames(t_df) <- c('Mean difference','CI lower','CI upper', 'Adjusted p value')

t_df
t_df$SE <- t_emm$contrasts.SE
t_df <- t_df[, c("Mean difference", "SE",'CI lower','CI upper', 'Adjusted p value')]
t_df


```


Make nice table of results
```{r}
p_load(knitr, kableExtra)
t_df %>% kable("html", digits=4, caption = "<center>Open Arm Time Percentage</center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, html_font = "Times New Roman")

```

```{r}
anova_light_pref <- aov(df$time_light ~ Genotype, data = df)
summary(anova_light_pref)
TukeyHSD(anova_light_pref)
#Personally, I prefer using the agricolae package which runs the command HSD.test. Use #'"unbalanced=TRUE" within the command to change the test to a Tukey-Kramer test.
p_load(agricolae)
tukey_result <- HSD.test(anova_light_pref, "Genotype", group = TRUE)
tukey_result
```

comparing relative darkness preference FREQUENCY

#assumption tests

```{r}
p_load(car)

#Independence
#Should be okay

#outliers- 0
p_load(rstatix)
outliers <- df %>%
  group_by(Genotype) %>%
  identify_outliers(light_preference_freq)

#normality - holds
by(df$light_preference_freq, df$Genotype, shapiro.test)

#Homogeneity of variance - holds
leveneTest(df$light_preference_freq, df$Genotype, center = mean)
bartlett.test(df$light_preference_freq ~df$Genotype, data=df)

kruskal.test(light_preference_freq ~ Genotype, 
             data = df)
```


```{r}

anova_light_pref <- aov(light_preference_freq ~ Genotype, data = df) #if using light_preference_freq instead, it is non-significant
summary(anova_light_pref)

#p_load(emmeans)
emm<-emmeans(anova_light_pref, pairwise ~ Genotype, conf.level = 0.95)
emm
t_emm<- as.data.frame(test(emm)[2])

Tukey_light <- TukeyHSD(anova_light_pref, conf.level = 0.95)

t_df<- as.data.frame(Tukey_light[1])
colnames(t_df) <- c('Mean difference','CI lower','CI upper', 'Adjusted p value')

t_df
t_df$SE <- t_emm$contrasts.SE
t_df <- t_df[, c("Mean difference", "SE",'CI lower','CI upper', 'Adjusted p value')]

```

Make nice table of results
```{r}
p_load(knitr, kableExtra)
t_df %>% kable("html", digits=4, caption = "<center>Open Arm Entry Percentage</center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, html_font = "Times New Roman")

```


Plots

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice
```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```


Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```


lightpreference time
```{r}
p_load(multcomp, broom, ggsignif)

Dunnet<- glht(anova_light_pref_t, mcp(Genotype = "Tukey")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df, aes(x=Genotype, y=light_preference)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  theme(text = element_text(family = "Times New Roman"))+
  ylab("Open Arm Time Percentage")+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(51,49,47,45,43,41)
  )

```

lightpreference frequency
```{r}

Dunnet<- glht(anova_light_pref, mcp(Genotype = "Tukey")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))


ggplot(data= df, aes(x=Genotype, y=light_preference_freq)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  theme(text = element_text(family = "Times New Roman"))+
  ylab("Open Arm Entry Percentage")+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(81,77,73,69,65,61)
  )

```
