---
title: "NOR"
output:
  pdf_document: default
  html_document: default
date: '2022-12-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}
setwd("~/Documents/TDP control files/Control study/")
# Loading
library(pacman)
p_load(readxl, tidyverse)
p_load(broom, ggsignif, multcomp, emmeans, DescTools)
# xlsx files
df <- read_excel("NOR_controls.xlsx")

```


Assign genotypes
```{r}
df$Genotype <- 0



df$Genotype <- ifelse(df$ID==82, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==84, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==107, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==109, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==80, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==86, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==115, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==117, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==74, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==76, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==119, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==121, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==129, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==137, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==141, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==98, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==102, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==112, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==120, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==124, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==128, 'NEFH-tTAxTDP', df$Genotype)

df$Genotype <- factor(df$Genotype, levels = c("WT", "NEFH-tTA", "TDP", "NEFH-tTAxTDP"))


```

Divide dataset up into phases with identical objects, and novel object

```{r}
Identical_cond <-  df %>% filter(Condition == "Hab")
Novel_cond<-  df %>% filter(Condition == "NovRGHT"| Condition =="NovLFT")

```

## Novel object condition

Calculating preference index for novel object over familiar object

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3332351/ 

In Wang et al.’s research (2007), they applied a measure of cognitive function through the Preference Index. This is a ratio of the amount of time spent exploring any one of the two objects in training phase (A or B) or the novel one in test phase (C) over the total time spent exploring both objects, i.e., A or B/(B + A) * 100 (%) in the training session and B or C/(B + C) * 100 (%) in the test phase. Therefore, a preference index above 50% indicates novel object preference, below 50% familiar object preference, and 50% no preference (Hammond et al. 2004).

```{r}
Novel_left <-  df %>% filter(Condition =="NovLFT")
Novel_right<-  df %>% filter(Condition == "NovRGHT")


# calculating DR_ index as (time novel  / (time novel + time familiar)*100)

Novel_right<- Novel_right %>% mutate(DR_index=((Novel_right$`In zone ContactZone_right  cumulative s`/(Novel_right$`In zone ContactZone_right  cumulative s`+Novel_right$`In zone ContactZone_left cumulative s`))*100))
Novel_left<- Novel_left %>% mutate(DR_index=((Novel_left$`In zone ContactZone_left cumulative s`/(Novel_left$`In zone ContactZone_left cumulative s`+Novel_left$`In zone ContactZone_right  cumulative s`))*100))

#DR index cannot be computed for M788 because it did not intercat with either object at all.

df2 <- rbind(Novel_left,Novel_right)

```

Dunnet test

```{r}

anova_novelty <- aov(df2$DR_index ~ Genotype, data = df2)
summary(anova_novelty)

#Dunnet's post hoc
DunnettTest(x=df2$DR_index, g=df2$Genotype)

```


Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice
```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```

Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```

```{r}

Dunnet<- glht(anova_novelty, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df2, aes(x=Genotype, y=DR_index)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Times New Roman"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Novelty Preference Index")+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(95,90,85)
  )

```

Tukey test

#assumption tests

```{r}
p_load(car)

#Independence
#Should be okay

#outliers - 2 outlier, but not extreme
p_load(rstatix)
outliers <- df2 %>%
  group_by(Genotype) %>%
  identify_outliers(DR_index)

#normality - holds
by(df2$DR_index, df2$Genotype, shapiro.test)

#Homogeneity of variance - holds
leveneTest(df2$DR_index, df2$Genotype, center = mean)

kruskal.test(DR_index ~ Genotype, 
             data = df2)
```


Checking preference for novel object over known object

```{r}

anova_novelty <- aov(df2$DR_index ~ Genotype, data = df2)
summary(anova_novelty)

#Dunnet's post hoc
p_load(emmeans)
emmeans(anova_novelty ,specs = trt.vs.ctrl~ Genotype)

#Tukey
TukeyHSD(anova_novelty , conf.level = 0.95)

emm<-emmeans(anova_novelty, pairwise ~ Genotype, conf.level = 0.95)
emm
n_emm<- as.data.frame(test(emm)[2])


Tukey_novelty <- TukeyHSD(anova_novelty, conf.level = 0.95)
n_df<- as.data.frame(Tukey_novelty[1])
colnames(n_df) <- c('Mean difference','CI lower','CI upper', 'Adjusted p value')

n_df$SE <- n_emm$contrasts.SE
n_df <- n_df[, c("Mean difference", "SE",'CI lower','CI upper', 'Adjusted p value')]

```

Make nice table of results
```{r}
p_load(knitr, kableExtra)
n_df %>% kable("html", digits=4, caption = "<center>Novelty Preference Index</center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, html_font = "Times New Roman")

```

```{r}

Dunnet<- glht(anova_novelty, mcp(Genotype = "Tukey")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df2, aes(x=Genotype, y=DR_index)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Times New Roman"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Novelty Preference Index")+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(110,105,100,95,90,85)
  )
```