---
title: "OFT"
output:
  pdf_document: default
  html_document: default
date: '2022-12-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load data
```{r}
setwd("~/Documents/TDP control files/Control study/")
# Loading
library(pacman)
p_load(readxl, tidyverse)
p_load(broom, ggsignif, multcomp, emmeans, DescTools)
# xlsx files
df <- read_excel("OFT_controls.xlsx")

```

Assign genotypes
```{r}
df$Genotype <- 0

df$Genotype <- ifelse(df$ID==82, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==84, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==107, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==109, 'TDP', df$Genotype)
df$Genotype <- ifelse(df$ID==80, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==86, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==115, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==117, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==74, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==76, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==119, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==121, 'NEFH-tTA', df$Genotype)
df$Genotype <- ifelse(df$ID==129, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==137, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==141, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==98, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==102, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==112, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==120, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==124, 'NEFH-tTAxTDP', df$Genotype)
df$Genotype <- ifelse(df$ID==128, 'NEFH-tTAxTDP', df$Genotype)

df$Genotype <- factor(df$Genotype, levels = c("WT", "NEFH-tTA", "TDP", "NEFH-tTAxTDP"))


```

We will look at two parameters:
- how much time mouse spent moving (absolute and relative to total time) (measure of exploration)
- how much time was spent on center compared to periphery (measure of anxiety, if anxious = more time spent in periphery)

Time spent moving:
```{r}
# divide by time not spent moving
df<-  df %>% mutate(relative_time_moving=(df$`Movement cumulative s`/(df$`Time not moving cumulative s`+df$`Movement cumulative s`))*100)
```

Periphery preference:

```{r}
df<-  df %>% mutate(periphery_preference=(`In zone periphery cumulative s`/(`In zone Center cumulative s`+`In zone periphery cumulative s`)*100))
```


Testing by genotype:


comparing distance moved

Used in  https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4354627/pdf/jove-96-52434.pdf paper from methods
```{r}
df$distance <- df$`Distance moved cm`

anova_distance_moved <- aov(df$distance ~ Genotype, data = df)
summary(anova_distance_moved )

DunnettTest(x=df$distance, g=df$Genotype)

```

Comparing periphery preference


```{r}

anova_periphery_preference <- aov(periphery_preference ~ Genotype, data = df)
summary(anova_periphery_preference)

DunnettTest(x=df$periphery_preference, g=df$Genotype)

```


 

Plots

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice
```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```


Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```

Periphery preference

```{r}

Dunnet<- glht(anova_periphery_preference, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df, aes(x=Genotype, y=periphery_preference)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Times New Roman"))+
  ylab("Periphery Time Percentage")+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(105,100,95)
  )
```


Distance moved
```{r}

Dunnet<- glht(anova_distance_moved, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df, aes(x=Genotype, y=distance)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  ylab("Distance Traveled (cm)")+
  theme(text = element_text(family = "Times New Roman"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(7000,6500,6000)
  )
```

Tukey tests

Distance moved
#assumption tests

```{r}
p_load(car)

#Independence
#Should be okay

#outliers - 1
p_load(rstatix)
outliers <- df %>%
  group_by(Genotype) %>%
  identify_outliers(`Distance moved cm`)

#normality - holds
by(df$`Distance moved cm`, df$Genotype, shapiro.test)

#Homogeneity of variance - does not hold
leveneTest(df$`Distance moved cm`, df$Genotype, center = mean)
bartlett.test(distance ~Genotype, df)

kruskal.test(df$`Distance moved cm` ~ Genotype, 
             data = df)
p_load(FSA)

dunnTest(distance ~ Genotype, 
             data = df,
         method="bonferroni")
```



```{r}

anova_distance_moved <- aov(df$distance ~ Genotype, data = df)
summary(anova_distance_moved )

TukeyHSD(anova_distance_moved , conf.level = 0.95)
emmeans(anova_distance_moved,specs = trt.vs.ctrl~ Genotype)

emm<-emmeans(anova_distance_moved, pairwise ~ Genotype, conf.level = 0.95)
emm
d_emm<- as.data.frame(test(emm)[2])


anova_distance_moved <- TukeyHSD(anova_distance_moved, conf.level = 0.95)
d_df<- as.data.frame(anova_distance_moved[1])
colnames(d_df) <- c('Mean difference','CI lower','CI upper', 'Adjusted p value')

d_df$SE <- d_emm$contrasts.SE
d_df <- d_df[, c("Mean difference", "SE",'CI lower','CI upper', 'Adjusted p value')]


```

Make nice table of results
```{r}
p_load(knitr, kableExtra)
d_df %>% kable("html", digits=4, caption = "<center>Distance Traveled (cm)</center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, html_font = "Times New Roman")

```

Periphery preference

#assumption tests

```{r}
p_load(car)

#Independence
#Should be okay

#outliers - 3
p_load(rstatix)
outliers <- df %>%
  group_by(Genotype) %>%
  identify_outliers(periphery_preference)

#normality - holds
by(df$periphery_preference, df$Genotype, shapiro.test)

#Homogeneity of variance - holds
leveneTest(df$periphery_preference, df$Genotype, center = mean)

kruskal.test(periphery_preference ~ Genotype, 
             data = df)
```



```{r}

anova_periphery_preference <- aov(periphery_preference ~ Genotype, data = df)
summary(anova_periphery_preference)

emmeans(anova_periphery_preference,specs = trt.vs.ctrl~ Genotype)
TukeyHSD(anova_periphery_preference, conf.level = 0.95)

emm<-emmeans(anova_periphery_preference, pairwise ~ Genotype, conf.level = 0.95)
emm
p_emm<- as.data.frame(test(emm)[2])


Tukey_periphery <- TukeyHSD(anova_periphery_preference, conf.level = 0.95)
p_df<- as.data.frame(Tukey_periphery[1])
colnames(p_df) <- c('Mean difference','CI lower','CI upper', 'Adjusted p value')

p_df$SE <- p_emm$contrasts.SE
p_df <- p_df[, c("Mean difference", "SE",'CI lower','CI upper', 'Adjusted p value')]


```


Make nice table of results
```{r}
p_load(knitr, kableExtra)
p_df %>% kable("html", digits=4, caption = "<center>Periphery Time Percentage</center>") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, html_font = "Times New Roman")

```

Plots

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice
```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```


Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```

Distance moved
```{r}
p_load(broom, ggsignif, multcomp)

anova_distance_moved <- aov(df$distance ~ Genotype, data = df)

Dunnet<- glht(anova_distance_moved, mcp(Genotype = "Tukey")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df, aes(x=Genotype, y=df$`Distance moved cm`)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  ylab("Distance Traveled (cm)")+
  theme(text = element_text(family = "Times New Roman"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(8500,8000,7500,7000,6500,6000)
  )
```
Periphery preference

```{r}

Dunnet<- glht(anova_periphery_preference, mcp(Genotype = "Tukey")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= df, aes(x=Genotype, y=periphery_preference)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Times New Roman"))+
  ylab("Periphery Time Percentage")+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Times New Roman",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(120,115,110,105,100,95)
  )
```

